Enabling Ambient Backscatter \\Using a Low-Cost Software Defined Radio

### 2017-04-07 paper.tex
Changed some subscripts to non-italic. Always use \text{}.

The received signal can be interpreted as
\begin{equation}
	s_{\text{IF}}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
\end{equation}     
This is multiplied with a cosine of the carrier frequency from a local oscillator.
\begin{multline}
        s_{\text{IF}}(t) \cdot s_\text{L}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
\end{multline}
With \ensuremath{2\,cos(a)cos(b)=cos(a-b)+cos(a+b)} and \ensuremath{2\,sin(a)cos(b)=sin(a+b)+sin(a-b)} follows
\begin{multline}
        2 \cdot s_{\text{IF}}(t) \cdot s_L(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
\end{multline}
One can see, that the interesting in-phase part (I) in this case is
represented by a DC value after mixing. With a low-pass filter this DC
value can be separated from the undesired rest. An analogous procedure
is done with the quadrature part (Q) when mixing with a sine.

@@

\subsection{Transmitter}
\begin{figure}[h]
\centering
\includegraphics[width=0.6\columnwidth]{./fig/s11v2}
\caption{Input reflection coefficient \ensuremath{\Gamma_{in}} also known as \ensuremath{S_{\text{11}}} of the self-made backscattering ground-plane antenna.}
\label{fig:s11}
\end{figure}

Our backscattering antenna is optimized for usage with the television signal,
which has its center frequency at 626 MHz. It is a ground plane antenna, which
has been built up on a large unetched PCB plate. A hole in the middle of the
PCB serves for connecting a wire to a 50 \ensuremath{\Omega} micro strip
transmission line. The wire, which acts as a vertical antenna element is
soldered to the transmission line. The microstrip transmission line design
equations given in \cite{pozar_microwave_2011} are implemented in various
programs e.g.  KiCAD, which has been used for designing them. An open-stub,
attached directly at the juncture where the vertical antenna element is
attached to the transmission line, is used for fine tuning of the input
impedance (cf. chapter 5 in \cite{pozar_microwave_2011}).  Accepting -15 dB as
sufficient value for \ensuremath{S_{11}}, one can say, that the antenna works
within a range from 626 MHz to 645 MHz. With our narrowband antenna the
undesired backscattering of out-of-band signals is reduced. With our approach
we were able to achieve an input impedance at 634 MHz of
\ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}. At 626 MHz
\ensuremath{S_{\text{11}}} is still approximately -17 dB.

@@

\begin{equation}
	\label{eq:awgn}
	C=B \cdot log_2(1+\frac{S}{N})= 25\,kHz \cdot log_2(11) \approx 86.5\,kbit/s
\end{equation}     
To do this it is important to use an anti-aliasing filter, as the Shannon-Nyquist theorem has to be satisfied: 
\begin{equation}
	\label{eq:nyquist}
	f_{samp} \geq 2 \cdot f_{a}
\end{equation}
\ensuremath{f_{a}} is the highest frequency in the signal. 
 A FIR filter
naively implemented with floats has been used to achieve this.  After
downsampling another filter is used to suppress noise.  The last demodulation
step is to rectify the signal (cf. equation \ref{eq:rectify}) and decide with a
software-defined Schmitt trigger whether a sample is a 1 or a 0.  
\begin{equation}
	\label{eq:rectify}
	abs(I,Q) = \sqrt{I^2 +  Q^2}
\end{equation} 
A registered listener receives the processed samples, which are broadcasted by
the demodulator. This registered listener is the decoder. After the decoder
received the samples it decides when a frame is sent or when the channel is
idle.  The decoder also includes a function to correlate the received bitstream
with an expected pattern. Hence it is able to determine the bit error ratio
(BER). Furthermore the code we have written so far subsumes different
simulators for e.g. playing back recoreded data. Another handy utility, which
has been written in Octave, is a tool (an oscilloscope) to look at the data,
which is written into files by the demodulator (cf. figure \ref{fig:transmission}).
\begin{figure}[h]

>>>>

The received signal can be interpreted as
\begin{equation}
	s_{\text{IF}}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
\end{equation}     
This is multiplied with a cosine of the carrier frequency from a local oscillator.
\begin{multline}
        s_{\text{IF}}(t) \cdot s_{\text{L}}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
\end{multline}
With \ensuremath{2\,cos(a)cos(b)=cos(a-b)+cos(a+b)} and \ensuremath{2\,sin(a)cos(b)=sin(a+b)+sin(a-b)} follows
\begin{multline}
        2 \cdot s_{\text{IF}}(t) \cdot s_{\text{L}}(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
\end{multline}
One can see, that the interesting in-phase part (I) in this case is
represented by a DC value after mixing. With a low-pass filter this DC
value can be separated from the undesired rest. An analogous procedure
is done with the quadrature part (Q) when mixing with a sine.

@@

\subsection{Transmitter}
\begin{figure}[h]
\centering
\includegraphics[width=0.6\columnwidth]{./fig/s11v2}
\caption{Input reflection coefficient \ensuremath{\Gamma_{\text{in}}} also known as \ensuremath{S_{\text{11}}} of the self-made backscattering ground-plane antenna.}
\label{fig:s11}
\end{figure}

Our backscattering antenna is optimized for usage with the television signal,
which has its center frequency at 626 MHz. It is a ground plane antenna, which
has been built up on a large unetched PCB plate. A hole in the middle of the
PCB serves for connecting a wire to a 50 \ensuremath{\Omega} micro strip
transmission line. The wire, which acts as a vertical antenna element is
soldered to the transmission line. The microstrip transmission line design
equations given in \cite{pozar_microwave_2011} are implemented in various
programs e.g.  KiCAD, which has been used for designing them. An open-stub,
attached directly at the juncture where the vertical antenna element is
attached to the transmission line, is used for fine tuning of the input
impedance (cf. chapter 5 in \cite{pozar_microwave_2011}).  Accepting -15 dB as
sufficient value for \ensuremath{S_{11}}, one can say, that the antenna works
within a range from 626 MHz to 645 MHz. With our narrowband antenna the
undesired backscattering of out-of-band signals is reduced. With our approach
we were able to achieve an input impedance at 634 MHz of
\ensuremath{Z_{\text{in}}=(52.2 + j 4.3) \Omega}. At 626 MHz
\ensuremath{S_{\text{11}}} is still approximately -17 dB.

@@

\begin{equation}
	\label{eq:awgn}
	C=B \cdot log_2(1+\frac{S}{N})= 25\,kHz \cdot log_2(11) \approx 86.5\,kbit/s
\end{equation}     
To do this it is important to use an anti-aliasing filter, as the Shannon-Nyquist theorem has to be satisfied: 
\begin{equation}
	\label{eq:nyquist}
	f_{\text{samp}} \geq 2 \cdot f_{\text{a}}
\end{equation}
\ensuremath{f_{\text{a}}} is the highest frequency in the signal. 
 A FIR filter
naively implemented with floats has been used to achieve this.  After
downsampling another filter is used to suppress noise.  The last demodulation
step is to rectify the signal (cf. equation \ref{eq:rectify}) and decide with a
software-defined Schmitt trigger whether a sample is a 1 or a 0.  
\begin{equation}
	\label{eq:rectify}
	abs(I,Q) = \sqrt{I^2 +  Q^2}
\end{equation} 
A registered listener receives the processed samples, which are broadcasted by
the demodulator. This registered listener is the decoder. After the decoder
received the samples it decides when a frame is sent or when the channel is
idle.  The decoder also includes a function to correlate the received bitstream
with an expected pattern. Hence it is able to determine the bit error ratio
(BER). Furthermore the code we have written so far subsumes different
simulators for e.g. playing back recoreded data. Another handy utility, which
has been written in Octave, is a tool (an oscilloscope) to look at the data,
which is written into files by the demodulator (cf. figure \ref{fig:transmission}).
\begin{figure}[h]
